---
title: "Analysis of Post-COVID-19 Inflation in Ghana"
author: "Benjamin Ekow Attabra"
date: '`r Sys.Date()`'
output:
  pdf_document: default
  html_document: default
---

# Introduction

The COVID-19 pandemic led to severe economic disruptions worldwide, and Ghana was no exception. One of the major consequences of the pandemic has been a significant rise in inflation, driven by various factors such as disruptions in the supply chain, changes in consumer behavior, and rising global food and fuel prices.

This project analyzes post-COVID-19 inflation in Ghana by investigating how a key economic factor, specifically **GDP growth**, has been influenced by inflation since 2020. We will use both exploratory data analysis and statistical modeling to draw conclusions.

# Data Preparation

## Loading the Data

We start by loading two datasets: one with inflation rate data and GDP growth.

```{r load-packages, message=FALSE}
library(dplyr)
library(ggplot2)
library(readr)

# Load the datasets
economic_data <- read_csv("Ghana Economic Data.csv")
inflation_data <- read_csv("Ghana_inflation_rate.csv")

# Inspect the first few rows
head(economic_data)
head(inflation_data)
```
## Data Cleaning and Merging

We clean the data by removing missing values and merging the datasets based on a common key, such as Date.
```{r}
# Clean the inflation dataset
inflation_data_clean <- na.omit(inflation_data)  # Remove missing values

# Merge the inflation data with the economic data based on 'Date'
combined_data <- left_join(inflation_data_clean, economic_data, by = "Year")

# Inspect the combined dataset
head(combined_data)
summary(combined_data)
```
## Exploratory Data Analysis (EDA)

We begin by exploring the inflation trends and their relationship with GDP growths.
Inflation Rate Trend
```{r}
# Plot the inflation rate over time
ggplot(combined_data, aes(x = Year, y = rate)) +
  geom_line(color = "blue") +
  theme_minimal() +
  labs(title = "Post-COVID-19 Inflation Trends in Ghana (1965-2023)",
       x = "Year",
       y = "Inflation Rate (%)")
```
## Relationship Between Inflation and Key Variables

Let’s explore the relationship between inflation and GDP growth
```{r}
# Plot relationship between GDP growth and inflation
ggplot(combined_data, aes(x = GDP_growth, y = rate)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red") +
  theme_minimal() +
  labs(title = "Impact of GDP Growth on Inflation",
       x = "GDP Growth (%)",
       y = "Inflation Rate (%)")
```

# Hypothesis Testing and Statistical Analysis

To test the significance of GDP growth on post-COVID-19 inflation, we perform hypothesis testing.

## Hypotheses

  - **Null hypothesis (H0):**There is no significant relationship between GDP growth and inflation rates.
  \[
  H_0: \mu = 0
  \]
- **Alternative hypothesis (H1):**There is a significant relationship between GDP growth and inflation rates.
  \[
  H_1: \mu \neq 0
  \]

 
 
We will perform a linear regression analysis to test this hypothesis.

## Regression Model
```{r}
# Fit a linear regression model
model <- lm(rate ~ GDP_growth, data = combined_data)

# Display the model summary
summary(model)
```
# Hypothesis Testing Results
```{r}
# Extract the p-value for GDP_growth
gdp_p_value <- summary(model)$coefficients["GDP_growth", "Pr(>|t|)"]

# Report the decision based on the p-value
if (gdp_p_value < 0.05) {
  decision <- "reject"
  conclusion <- "significant"
} else {
  decision <- "fail to reject"
  conclusion <- "no significant"
}

# Output the conclusion
cat("The p-value for GDP growth is ", gdp_p_value, ". Therefore,\nwe", decision, 
    " the null hypothesis and conclude that there is", conclusion, 
    " relationship \nbetween GDP growth and inflation post-COVID-19.")
```
# Conclusion

This analysis of post-COVID-19 inflation in Ghana demonstrates how GDP growth has influenced inflation. Based on our hypothesis testing, We found insufficient evidence to support a significant relationship between GDP growth and inflation in Ghana. The insights from this analysis could be valuable for policymakers as they craft strategies for economic recovery.

# References

    World Bank. "Ghana Economic Outlook."
    IMF. "Post-COVID-19 Recovery and Inflation Reports."

